\chapter{Bounding Techniques for DPOR}
\label{bounded}

In this chapter Bounding Techniques for DPOR are discussed as well as the challenges emerging from these techniques.
The challenges of bounding the DPOR have already been discussed.

%% Challenges discussion can not be placed before the presentation of the algorithms since it is closely related to the algorithms themselves
%% and not the Bounding of DPOR. Provided that other algorithms were suggested, different challenges may had been arisen. 

\section{Naive-BPOR}

The first bounded technique to be presented is the Naive-BPOR (Algorithm \ref{Vanilla}). The purpose of the algorithm is to block traces that exceed the bound
limit. 

\begin{algorithm}
    \caption{Naive-BPOR}
    \label{Vanilla}
    \SetKwInOut{Input}{input}
    Explore($\langle \rangle$,$\emptyset$,$b$)\;
    \Fn{Explore($E$,$Sleep$,$b$)}{
        \If{$\exists p \in (enabled(s_{[E]}) \backslash Sleep)$ such that $B_v(E.p) \leq b)$ }{
            $backtrack(E) :={p}$ \;
            \While{$\exists p \in (backtrack(E) \backslash Sleep$ and $B_v(E.p) \leq b$}{
                \ForEach{$e \in dom(E)$ such that $e \lesssim_{E.p} next_{[E]}(p) $}{
                    \Let{$E' = pre(E,e)$}
                    \Let{$u = notdep(e,E).p$}
                    \If{$I_{E'}(u) \cap backtrack(E') = \emptyset$}{
                        add some $q' \in I_{[E']}(u) to backtrack(E') $ \;
                    }
                }
                \Let{$Sleep' := \{q \in Sleep \mid E \models p \diamondsuit q \} $}
                $Explore(E.p, Sleep, b)$ \;
                add $p$ to $Sleep$ \;

            }
        }
    }
\end{algorithm}

The Algorithm \ref{Vanilla} is almost the same with Source-DPOR(Algorithm \ref{Source}). The only additions made are
related to the thread scheduling. When a a step of a process $p$ added to $E$ result the trace $E.p : B_v(E.p) > b$ then
this process is not allowed to be scheduled. This algorithm is not sound i.e., it does not examine every trace that
compensates with the bound

Let the writer-2-readers example with $b=0$. An example of exploration for bound $b=0$ is given in Figure \ref{Naive-BPOR for bound=$0$}. 

\label{Vanilla0}    
\trace{w2rvbound.pdf}{Naive-BPOR for bound=$0$}

As we can see there are 4 traces that do not exceed the bound. These are:
$p.q.q.r.r$, $q.q.p.r.r$, $r.r.p.q.q$, $q.q.r.r.p$.
However the Naive-BPOR is not able to explore them all; $r.r.p.q.q$ is not explored.
As it was shown in the comparison of persistent and source sets, $r$ is never registered as the first event of the trace
since this will lead to a sleep set blocked trace. The branch that would lead to an equivalent trace to $r.r.p.q.q$ is rejected
since it would have higher bound count. Interestingly enough a Naive-BPOR based on persistent sets would explore a larger state-space since
it would have explored would probably have explored the trace $r.r.p.q.q$.

\section{BPOR}

The next step is to implement a sound preemption bounded algorithm (BPOR) \cite{BPOR}. This algorithm would rely on
persistent sets. Since we have correctly implemented persistent sets for Nidhugg-DPOR we can easily implement a BPOR
algorithm. The novelty of the BPOR is the introduction of conservative branches. These are branches that are introduced
in order to guarantee the exploration of the whole state space. It is common for a trace to exceed the bound limit
whereas there is an equivalent trace which does not. The conservative branches are used for this purpose.

To make this idea more clear we introduce the notion of Trace block.
\begin{definition}{(Trace block)}
For a trace $T$ a sequence $B$ of consecutive events is a trace block iff all events happen in the same thread i.e. all
the events have the same thread id.
\end{definition}

The idea behind conservative branches is quit simple. When a branch is added a conservative branch is added at the
beginning of the corresponding block. Usually concurrent events take place inside a block. As a result when a branch is
taken then the preemption count will most probably increased. However, had this branch been added at the beginning of
the block the preemption count would not have been increased. BPOR is presented in detail in Algorithm \ref{BPOR}. The
BPOR algorithm differs from Classic-DPOR in the double nested loop which is introduced in line 3. This inner loop is
introduced in order for the BPOR to compensate with the defintion of the preemption-bounded persistent sets which was
introduced in Chapter 2. Moreover, as in Naive-BPOR, the exploration is also be terminated if a trace exceeds the bound
limit.

\begin{algorithm}
    \caption{BPOR}
    \label{BPOR}
    \SetKwInOut{Input}{input}
    \SetKwInput{Initialization}{Explore($ \emptyset $)}
    \SetKwHangingKw{Let}{let}
    \Fn{Explore($E$)}{
        \Let{$s := last(E)$}
        \For{all process $p$}{
            \For{all process $q \neq p$}{
            \If{$\exists i = max(\{ i \in dom(E) \mid E_i$ is dependent and may be co-enabled with $next(s,p)$ and $ E_i.tid = q \} $}{
                \uIf{$p \in enabled(pre(E,i)))$}{
                    add $p$ to $backtrack(pre(E,i))$ \;
                }
                \Else{add $enabled(pre(E,i))$ to $backtrack(pre(E,i))$ \;}
                \uIf{$j = max(\{ j \in dom(E) \mid j = 0 $ or $ S_{j-1}.tid \neq S_j.tid $ and $ j<i \})$}{
                    \uIf{$p \in enabled(pre(E,i)))$}{
                        add $p$ to $backtrack(pre(E,i))$ \;
                    }
                    \Else{add $enabled(pre(E,i))$ to $backtrack(pre(E,i))$ \;}
                }
            }
            }
        }
        \If{$p \in enabled(s)$}{
            add $p$ to $backtrack(s)$ \;
        }
        \Else{
            add any $u \in enabled(s)$ to $backtrack(s)$ \;
        }
        \Let{$visited = \emptyset $}
        \While{$ \exists u \in (enabled(s) \cap backtrack(s) \backslash visited) $}{
            add $u$ to $visited$ \;
            \uIf{$(B_v(S.next(s, u)) \leq c)$}{
                Explore($S.next(s, u)$) \;
            }
        }
    }
\end{algorithm}

\section{Nidhugg-BPOR}

In order to take advantage of Nidhuggs's infrastructure, a modification of this algorithm is used.
The algorithm is presented in Algorithm \ref{Nidhugg BPOR}.

\begin{algorithm}
    \caption{Nidhugg-BPOR}
    \label{Nidhugg BPOR}
    \SetKwInOut{Input}{input}
    \SetKwHangingKw{Let}{let}
    Explore($\langle \rangle$,$\emptyset$,$b$)\;
    \Fn{Explore($E$,$Sleep$,$b$)}{
        \If{$\exists p \in ((enabled(s_{[E]}) \backslash Sleep)$ and $B_v(E.p) <=b$}{
            backtrack(E) $:={p}$ \;
            \While{$\exists p \in (backtrack(E) \backslash Sleep $ and $B_v(E.p) <=b$}{
                \ForEach{$e \in dom(E)$ such that $e \lesssim_{E.p} next_{[E]}(p)$}{
                    \Let{$E' = pre(E,e)$}
                    \Let{$u = notdep(e,E).p$}
                    \Let{$CI = \{ i \in I_{E'}(u) \mid i \rightarrow p \}$}
                    \If{$CI \cap backtrack(E') = \emptyset$}{
                        \If{$CI \neq \emptyset$}{
                            add some $q' \in CI$ to $backtrack(E') $ \;
                        }
                        \Else{
                             add some $q' \in I_{[E']}(u)$ to $backtrack(E') $ \;}
                        }
                    \Let{$E''= pre\_block(e,E)$}
                    \Let{$u = notdep(e,E).p$}
                    \Let{$CI = \{ i \in I_{E''}(u) \mid i \rightarrow p \}$}
                    \If{$CI \cap backtrack(E') = \emptyset$}{
                        \If{$CI \neq \emptyset$}{
                            add some $q' \in CI$ to $backtrack(E') $ \;
                        }
                        \Else{
                            add some $c(q') \in I_{[E'']}(u)$ to $backtrack(E'') $ \;
                        }
                    } 
                }
                \Let{ $Sleep' := \{q \in Sleep \mid E \models p \diamondsuit q \}$ } 
                $Explore(E.p, Sleep)$ \;
                \If{$p$ is not conservative}{
                    add $p$ to $Sleep$ \;
                }
            }
        }
    }
\end{algorithm}

Notice that persistent sets are calculated for both the conservative and the non-conservative branches.

A critical challenge arises when a DPOR algorithm is used in tandem with sleep sets. This stems from the fact that
conservative branches are not added due to a concurrent event. By observing the sleep set algorithm we notice that if we
follow the same strategy as with non-conservative branches many traces will end up being blocked.

Let us take the writer-2readers example as shown in Figure \ref{Usage of non-conservative branches}.

\trace{w2rpersistent.pdf}{Usage of non-conservative branches}

We notice that the last trace is sleep set blocked while it should be examined. The algorithm is unaware that the
process $r$ should be removed from the sleep set since there is no or will ever be found any conflict with the first
command of the thread which is related with a non shared variable. In order to deal with this problem, when a
conservative branch is chosen then it should not be added to the sleep set. However there must be a set recording all
the branches that were added at this certain point of the trace so no thread is added twice. Without this set the
exploration would continue indefinitely. The solution is based on the notion of the conservative sets where every thread
that was added to the branch is recorded. 

Intuitively the algorithm is the same with the Nidhugg-DPOR with the addition of the conservative branches. 
The algorithm is based on the idea of the conservative sets where every thread that was added to the branch is recorded.  
However many challenges arise which are discussed in the implementation section.

\tracelong{w2rbpor.pdf}{Example of BPOR execution}

\section{Source-BPOR}

Having discussed BPOR algorithm and Nidhugg-BPOR, the next step is to try combine source sets with the algorithm. The
first observation we have to make is that source sets and thus the algorithm for creating these sets is not suitable for
adding conservative branches. A quick explanation is given in the next writer-2-readers example even though the problem
will be further discussed later. Let us assume that we have followed the source set algorithm for adding conservative
sets. The results are shown at Figure \ref{Following source sets for conservative branches}.

\trace{w2rsourceconservative.pdf}{Following source sets for conservative branches}

It is clear that some traces are not explored. Specifically, the trace which start with r has been rejected. The reason
is that it shares the same initials with r1 even at the beginning of that block. As a result the algorithm must create
to persistent sets when conservative branches are added. 

Having made the preceding observations the algorithm used for Source-BPOR is shown in Algorithm \ref{SBPOR}.

\begin{algorithm}
    \caption{Source-BPOR}
    \label{SBPOR}
    \SetKwInOut{Input}{input}
    \SetKwHangingKw{Let}{let}
    Explore($\langle \rangle$,$\emptyset$,$b$)\;
    \Fn{Explore($E$,$Sleep$,$b$)}{
        \If{$\exists p \in ((enabled(s_{[E]}) \backslash Sleep)$ and $B_v(E.p) <=b$}{
            backtrack(E) $:={p}$ \;
            \While{$\exists p \in (backtrack(E) \backslash Sleep $ and $B_v(E.p) <=b$}{
                \ForEach{$e \in dom(E)$ such that $e \lesssim_{E.p} next_{[E]}(p)$}{
                    \Let{$E' = pre(E,e)$}
                    \Let{$u = notdep(e,E).p$}
                    \If{$I_{E'}(u) \cap backtrack(E') = \emptyset$}{
                        add some $q' \in I_{[E']}(u) \text{ to } backtrack(E') $ \;
                    }
                    \Let{$E''= pre\_block(e,E)$}
                    \Let{$u = notdep(e,E).p$}
                    \Let{$CI = \{ i \in I_{E''}(u) \mid i \rightarrow p \}$}
                    \If{$CI \cap backtrack(E') = \emptyset$}{
                        \If{$CI \neq \emptyset$}{
                            add some $q' \in CI$ to $backtrack(E') $ \;
                        }
                        \Else{
                            add some $c(q') \in I_{[E'']}(u)$ to $backtrack(E'') $ \;
                        }
                    } 
                }
                \Let{ $Sleep' := \{q \in Sleep \mid E \models p \diamondsuit q \}$ } 
                $Explore(E.p, Sleep)$ \;
                \If{$p$ is not conservative}{
                    add $p$ to $Sleep$ \;
                }
            }
        }
    }
\end{algorithm}

We can notice that the Algorithm \ref{SBPOR} works as Source-DPOR (Alg. \ref{Source}) for non-conservative branches and as
BPOR (Alg. \ref{Nidhugg BPOR}) for conservative branches .

\section{Challenges Arising from the addition of conservative branches}
%% I still believe that this discussion should be placed after results have been presented.

\subsection{Conservative Branches}
In a previous section we saw that conservative sets cannot utilize the sleep set optimization. This is due
to the fact that these branches are not produced by conflicts and as a result it is impossible to ``wake up'' another
process whose next step may be a local operation. The problem is getting even more complex considering that when a
conservative branch is added the algorithm ``forgets'' what was previously in this place. This lack of memory leads to an
explosion of the state space. This explosion is greater than the explosion happening when exploring the complete state
space. In order to explain this better an example involving a writer and three readers is given in Figure \ref{writer-3-readers
explosion}. Specifically, the writer process just writes a value on a variable x while the reader processes read this variable
after reading another variable unique for each process which can be regarded as a local operation.

\trace{w3rbpor.pdf}{writer-3-readers explosion}

As we can infer from the example there are more states to be explored than the expected ones. The expected number of
traces would be $8$ but $4!$ are explored. Notice that every reader is backtracked in the first command since it
conflicts with $w(x)$. These branches should be blocked by the sleep sets in the unbounded search. However, these
branches are also conservative ones. There are extra traces such as the $r2.r1.w$ trace which explores the already
explored $r1.r2.w$ trace. The algorithm is not aware when adding the conservative trace whether an equivalent trace will
be explored. The situation is worse when more readers are to read the same global variable. The total traces explored
for a big bound approach the $N!$ where $N$ the total number of threads. We can easily notice that execution seems to be
aware of the previous executions since sleep sets cannot be used. This happens due to the redundant inversions of the
reading operations. 

\subsection{Sleep Sets}
The results of the various bounding algorithms suggest that the number of sleep set blocked traces is really small
compared to the number of the explored traces. This is due to the conservative branches. It was made clear that when
both a conservative and a non-conservative branch of the same thread is added then the conservative branch prevails.
This trace would be redundant in an unbounded version of the algorithm but it is not in a bounded version since the
non-conservative branch may had been rejected. Moreover, even if both threads had been accepted by the algorithm there
may be a later scheduling which may be rejected when the trace was caused by the non-conservative branch and be accepted
by the equivalent trace caused by the conservative branch. 

Another ``problem'' with the sleep sets is that are ``in favor'' of the branches that increase the bound count while
they block traces with lower bound count. In Figure \ref{Sleep set contradiction} it is shown that the branch which
would not have caused a increase of the bound is rejected.

\trace{sleepsetproblem.pdf}{Sleep set contradiction}

This behavior is quite reasonable if we consider the depth-first nature of the algorithm. As it was demonstrated in the
first chapter, the DPOR algorithm performs a DFS-search. As a result it takes the branches which are located lower in
the traces where the bound count is higher since more preemptive switches have taken place. The purpose of the sleep
sets is to block redundant traces, i.e. traces that have already been explored. As a result the equivalent traces with
lower bound count will be rejected. A method that would explore the state space in breadth first way would be unfeasible
since it would entail a huge memory overhead which stems from the storage of the traces that have not be completely
explored. 

\iffalse
\subsection{Source Sets - Optimal DPOR}
The source set technique which can lead many times to optimal coverage of the state space manages avoid the exploration
of redundant sets. However as it was discussed in the previous section it avoids the scheduling of the traces.
Unfortunately it cannot be used in the when conservative branches are added since these branches are not related to the
sleep sets. As a result the conservative branches alone will never lead to sleep set blocked traces.

However, in many test cases there are some sleep set blocked traces which are caused by conditional reads and writes.
When a technique which does not utilize the sleep sets is to be used these traces would have been easily eliminated.
Unfortunately, it was experimentally shown that explored traces outnumber the sleepset blocked traces and, consequently,
the implementation of such an algorithm would have a minor impact.

Moreover, we have shown that even if we maintain the source-set optimization for the non-conservative branches the
results will be equivalent with using persistent sets. The idea of keeping the rejected traces from Source-DPOR that
would have been added from DPOR (with persistent sets) was rejected since it harms the soundness of the algorithm.
\fi

\subsection {The Performance - Soundness Tradeoff}

From the previous discussion it is clear that all the bounding algorithms have to deal with a tradeoff. Some algorithms may be faster but
explore a smaller amount of traces, without even covering the full state space within the bound (Algorithm \ref{Vanilla}), 
whereas others explore the whole state space (Algorithm \ref{Nidhugg BPOR}) within the bound but require more time.
